\mytitle

### 2021-12-20 paper.tex
extend conclusions
move footnote down

In the following an ILP-aware implementation of the real dilogarithm
is presented, which allows for the parallel execution of different
floating\-/point operations.  The real dilogarithm is defined as
%
\begin{align}
  \Li(x) &= -\int_0^x \frac{\ln(1-t)}{t}dt, &
  &x<1.
\end{align}
%
For $|x|<1$ the real dilogarithm has the series representation
%
\begin{align}
  \Li(x) = \sum_{k=1}^\infty \frac{x^k}{k^2}.
  \label{eq:series}
\end{align}
%
Further series representations exist, for example the accelerated
series
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty \frac{B_k}{(k+1)!} [-\ln(1-x)]^{k+1},
\end{align}
%
where $B_k$ are the Bernoulli numbers.  A series representation in
terms of Chebyshev polynomials of the first kind $T_n(x)$ is given in
\cite{luke},
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty a_n T_n(2x-1),
  \label{eq:cheby}
\end{align}
%
which is used for example in the ROOT program library \cite{root}.
The r.h.s.\ of Eqs.~\eqref{eq:series}--\eqref{eq:cheby} can be summed,
for example, using Horner's scheme or Clenshaw's algorithm
\cite{clenshaw}.  These summation techniques are, however, purely
sequential and thus do not make use of ILP.  In the following an
approximation of the real dilogarithm is presented, which makes use of
ILP, using a rational function approximation in combination with
Estrin's scheme, similar to the approach used in
Eq.~\eqref{eq:mixed}.%
\footnote{The Cephes library \cite{cephes} implements an equivalent
  rational function approximation for Spence's function
  $\operatorname{Sp}(x)=\Li(1-x)$.}

@@

The third fastest implementation is provided by Cephes, which is by a
factor two slower than the implementation prestend in this paper.
This longer runtime has several reasons: For one, Cephes uses a
rational function approximation with more terms in the numerator and
denominator.  Furthermore, Cephes evaluates the numerator and
denominator polynomials using Horner's scheme, which is purely
sequential.\footnote{For $x>1$ Cephes calculates $\ln(x)$ twice, which
  leads to a further increased runtime when a benchmark interval with
  $x>1$ is chosen.}

@@

In this paper different approaches for the numerical calculation of
the real dilogarithm have been compared w.r.t.\ their potential for
the use of instruction\-/level parallelism.  A particular
implementation of a rational minimax approximant has been proposed and
compared to several existing implementations w.r.t.\ runtime.  C, C++
and FORTRAN implementations of the presented algorithm can be found in
the Polylogarithm library \cite{polylogarithm}.  Implementations in
Julia and Rust can be found in \cite{PolyLog.jl} and \cite{polylog},
respectively.

>>>>

In the following an ILP-aware implementation of the real dilogarithm
is presented, which allows for the parallel execution of different
floating\-/point operations.  The real dilogarithm is defined as
%
\begin{align}
  \Li(x) &= -\int_0^x \frac{\ln(1-t)}{t}dt, &
  &x<1.
\end{align}
%
For $|x|<1$ the real dilogarithm has the series representation
%
\begin{align}
  \Li(x) = \sum_{k=1}^\infty \frac{x^k}{k^2}.
  \label{eq:series}
\end{align}
%
Further series representations exist, for example the accelerated
series
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty \frac{B_k}{(k+1)!} [-\ln(1-x)]^{k+1},
\end{align}
%
where $B_k$ are the Bernoulli numbers.  A series representation in
terms of Chebyshev polynomials of the first kind $T_n(x)$ is given in
\cite{luke},
%
\begin{align}
  \Li(x) = \sum_{k=0}^\infty a_n T_n(2x-1),
  \label{eq:cheby}
\end{align}
%
which is used for example in the ROOT program library \cite{root}.
The r.h.s.\ of Eqs.~\eqref{eq:series}--\eqref{eq:cheby} can be summed,
for example, using Horner's scheme or Clenshaw's algorithm
\cite{clenshaw}.  These summation techniques are, however, purely
sequential and thus do not make use of ILP.  In the following an
approximation of the real dilogarithm is presented, which makes use of
ILP, using a rational function approximation in combination with
Estrin's scheme, similar to the approach used in
Eq.~\eqref{eq:mixed}.

@@

The third fastest implementation is provided by Cephes \cite{cephes},
which implements a rational function approximation for Spence's
function $\operatorname{Sp}(x)=\Li(1-x)$.  The implementation provided
by Cephes is by a factor two slower than the implementation prestend
in this paper.  This longer runtime has several reasons: For one,
Cephes uses a rational function approximation with more terms in the
numerator and denominator.  Furthermore, Cephes evaluates the
numerator and denominator polynomials using Horner's scheme, which is
purely sequential.\footnote{For $x>1$ Cephes calculates $\ln(x)$
  twice, which leads to a further increased runtime when a benchmark
  interval with $x>1$ is chosen.}

@@

In this paper different approaches for the numerical calculation of
the real dilogarithm have been compared w.r.t.\ their potential for
the use of instruction\-/level parallelism.  A particular
implementation of a rational minimax approximant has been proposed and
compared to several existing implementations w.r.t.\ runtime.  A C
implementations of the presented algorithm can be found in the
appendix.  C++ and FORTRAN implementations can be found in the
Polylogarithm library \cite{polylogarithm}.  Implementations in Julia
and Rust can be found in \cite{PolyLog.jl} and \cite{polylog},
respectively.
